---
title: "TauDEM analysis"
author: "Matthew Ross"
date: "May 30, 2017"
output:
  html_document:
    toc: true
    theme: united
editor_options: 
  chunk_output_type: console
---



# TauDEM

This is a script that does terrain analyses using the [TauDEM tool](http://hydrology.usu.edu/taudem/taudem5/index.html) from David Tarboton at Utah State. All scripts below should be run in bash or terminal after installing TauDEM. R Code is mixed throughout to generate a final data table for analysis. 

Some R code is at the bottom of the script to project outlets and subset rasters. 


## Pit remove
```{bash, eval=F}
mpiexec -n 8 pitremove -z data/in/TauNew.tif -fel data/out/newfel.tif

mpiexec -n 8 pitremove -z data/in/TauOldElev.tif -fel data/out/oldfel.tif
```

## D8 

### Flow direction with slope and flow direction
```{bash, eval=F}
mpiexec -n 8 d8flowdir -p data/out/newdir8.tif -sd8 data/out/newsd8.tif -fel data/out/newfel.tif

mpiexec -n 8 d8flowdir -p data/out/olddir8.tif -sd8 data/out/oldsd8.tif -fel data/out/oldfel.tif

```

### Flow accumulation
```{bash, eval=F}
mpiexec -n 8 aread8 -p data/out/newdir8.tif -ad8 data/out/newuaa8.tif

mpiexec -n 8 aread8 -p data/out/olddir8.tif -ad8 data/out/olduaa8.tif
```

 
##  D-Infinity

### Flowdir
```{bash, eval=F}
mpiexec -n 8 dinfflowdir -ang data/out/newinffdr.tif -slp data/out/newinfslp.tif -fel data/out/newfel.tif

mpiexec -n 8 dinfflowdir -ang data/out/oldinffdr.tif -slp data/out/oldinfslp.tif -fel data/out/oldfel.tif


```

### Flow accumulation
```{bash, eval=F}
mpiexec -n 7 areadinf -ang data/out/newinffdr.tif -sca data/out/newinfuaa.tif

mpiexec -n 7 areadinf -ang data/out/oldinffdr.tif -sca data/out/oldinfuaa.tif

```


##  UAA analysis

### Threshold 1000 pixels = 0.1 km2

```{bash}

#New infinity
mpiexec -n 7 threshold -ssa data/out/newinfuaa.tif -src data/out/new100inf.tif -thresh 100
#Old infinity 
mpiexec -n 7 threshold -ssa data/out/oldinfuaa.tif -src data/out/old100inf.tif -thresh 100

#New D8 
mpiexec -n 7 threshold -ssa data/out/newuaa8.tif -src data/out/new1008.tif -thresh 100
#Old D8
mpiexec -n 7 threshold -ssa data/out/olduaa8.tif -src data/out/old1008.tif -thresh 100

#Curvature method



```


### Delineate watersheds

```{bash, eval=F}
#New Infinity
mpiexec -n 7 areadinf -ang data/out/newinffdr.tif -o data/in/shapefile/approxnew.shp -sca newshedinf.tif

mpiexec -n 7 aread8 -p data/out/newdir8.tif -o data/in/shapefile/approxnew.shp -ad8 newshed8.tif

mpiexec -n 7 areadinf -ang data/out/oldinffdr.tif -o data/in/shapefile/approxnew.shp -sca oldshedinf.tif

mpiexec -n 7 aread8 -p data/out/olddir8.tif -o data/in/shapefile/approxnew.shp -ad8 oldshed8.tif

```



##  Stack and cutup rasters to smaller extents for later analysis in r. 

```{r, eval=F}
#Using a new faster raster library called Velox


#Setup readin paths and crop extents and name order and empty lists
#Upslop accumulated area data
uaa.paths <- paste0('data/out/',c('newshedinf.tif', 'oldshedinf.tif','newshed8.tif','oldshed8.tif'))
#Slope data. 
slope.paths <- paste0('data/out/',c('newinfslp.tif','oldinfslp.tif','newsd8.tif','oldsd8.tif'))
#Elevation data. 
e.paths <- paste0('data/in/',c('TauNew.tif','TauNew.tif','TauOldElev.tif','TauOldElev.tif')
#Names for where elevation profiles come from (matched with current shed boundaries or not? New.e.new means New elevatio profiles with new shed boundaries. New.e.old means new elevation with old shed boundaries)
e.names <- c('New.e.new','New.e.old','Old.e.new','Old.e.old')
#Read in raster outlines
sheds <- raster(uaa.paths[1]) %>% crop(.,extent(400000,470000,4150000,4230000))
oldsheds<- raster(uaa.paths[3]) %>% crop(.,extent(400000,470000,4150000,4230000))


#Set all values to 1
sheds[sheds > 0] <- 1



#Turn these into shapefiles
shd.shp <- gdal_polygonizeR(sheds)

#Buffer out by 500m to include any ridge changes and reproject to outlet projeciton
shd.shp.buf <- gBuffer(shd.shp,width=100) %>%
  spTransform(.,projection(outs)) %>%
  disaggregate(.) %>%
  SpatialPolygonsDataFrame(.,data=data.frame(id=1:10))


##Plot shd.shp.buf to see if the buffered area includes old watersheds
# plot(shd.shp.buf)
# plot(oldsheds,col='black',add=T)


#Join these shapefiles to the outlets to get the names (ref or not ref)
shd.names <- over(shd.shp.buf,outs)
shd.shp.names <- spCbind(shd.shp.buf,shd.names)


#save(shd.shp.names,file='data/out/watersheds.RData')

#Setup storage lists
uaa.stack <- list()
slope.stack <- list()
e.stack <- list()
esu.stacked <- list()
#Evaluate on 7 cores 
cl <- makeCluster(3)
registerDoParallel(cl,cores=3)
all.list <- foreach(i = 1:nrow(shd.names)) %dopar% {
  library(raster)
  library(rgdal)
  library(magrittr)
  #Read in each Watershed outline to trim data
  shape <- shd.shp.names[shd.shp.names$Name == shd.names$Name[i],]
  for(j in 1:length(uaa.paths)){
      #Read in watershed delineations and crop to single watershed
      shed <- raster(uaa.paths[j]) %>% crop(.,shape)
      #Store this cropped watershed uaa tif in a list
      uaa.stack[[j]] <- shed
      #Read in slope raster and crop
      slope <- raster(slope.paths[j]) %>% crop(.,shape)
      #Set Slope values outside of watershed to NA
      slope[is.na(shed)] <- NA
      #Store in slope stack
      slope.stack[[j]] <- slope
      #Samesies with elevation
      e <- raster(e.paths[j]) %>% crop(.,shape)
      e[is.na(shed)] <- NA
      e.stack[[j]] <- e
  }
  esu.stacked[[i]] <- stack(c(e.stack,slope.stack,uaa.stack)) %>% trim(.)
}
endCluster()
names(all.list) <- shd.names$Name


#save(all.list, file='data/out/ras.stack.ref.RData')
```



##  Convert raster data to data frames
```{r}
#Loads in all list which is a list of all sites
load('data/out/ras.stack.ref.RData')

library(doParallel) #to do parallel processing
library(parallel) # to register parallel processing
library(foreach) # to do parallel processing as a forloop
#Evaluate on 7 cores
cl <- makeCluster(3) #make a cluster with 3 cores
registerDoParallel(cl,cores=3) #tell parallel processor to use 3 cores

stack.list <- foreach(i=1:length(all.list)) %dopar% {
  library(raster)
  library(dplyr)
  library(magrittr)
  l <- extract(all.list[[i]]) %>% as.data.frame(.) %>%
    mutate(Site = shd.names$Name[i]) 
}
endCluster()

stack.df <- do.call('rbind',stack.list) 
names(stack.df)[1:4] <- e.names

  

head(stack.df)
#Name the rows of this wide data frame

save(stack.df,file='data/out/UAA.E.Slope.Ref.RData')


```







