
# Identifying heavily mined watersheds

The goal of othis project is to better understand how mining has 
altered the geomorphology of these landscapes, so we want to work in landscapes
that have been heavily mined. 

## Project outlets with same projection as rasters

```{r, eval=FALSE}

library(raster)
library(tidyverse)
library(magrittr)
library(rgdal)
library(rgeos)
library(shapefiles)
library(maptools)
library(leaflet)
library(velox)
library(foreach)
library(snow)
library(parallel)
library(doParallel)
library(Hmisc)

```





### Convert river network into shapefile

Note: All of this is using gdal/sp/etc... I would recommend using `sf`, `s2` and 
`terra` for future approaches. As stated at beginning this project was started
in 2014!

This using code from [John Baumgartner](https://johnbaumgartner.wordpress.com/2012/07/26/getting-rasters-into-shape-from-r/) note that gdal and python-gdal must be installed (on linux machine)

I used this network to hand select watershed outlets based on a separate layer of mining extent.
I tried to capture watersheds that were > 90% mined. 

```{r, eval=F}


gdal_polygonizeR <- function(x, outshape=NULL, gdalformat = 'ESRI Shapefile',
                             pypath=NULL, readpoly=TRUE, quiet=TRUE) {
  if (isTRUE(readpoly)) require(rgdal)
  if (is.null(pypath)) {
    pypath <- Sys.which('gdal_polygonize.py')
  }
  if (!file.exists(pypath)) stop("Can't find gdal_polygonize.py on your system.")
  owd <- getwd()
  on.exit(setwd(owd))
  setwd(dirname(pypath))
  if (!is.null(outshape)) {
    outshape <- sub('\\.shp$', '', outshape)
    f.exists <- file.exists(paste(outshape, c('shp', 'shx', 'dbf'), sep='.'))
    if (any(f.exists))
      stop(sprintf('File already exists: %s',
                   toString(paste(outshape, c('shp', 'shx', 'dbf'),
                                  sep='.')[f.exists])), call.=FALSE)
  } else outshape <- tempfile()
  if (is(x, 'Raster')) {
    require(raster)
    writeRaster(x, {f <- tempfile(fileext='.tif')})
    rastpath <- normalizePath(f)
  } else if (is.character(x)) {
    rastpath <- normalizePath(x)
  } else stop('x must be a file path (character string), or a Raster object.')
  system2('python', args=(sprintf('"%1$s" "%2$s" -f "%3$s" "%4$s.shp"',
                                  pypath, rastpath, gdalformat, outshape)))
  if (isTRUE(readpoly)) {
    shp <- readOGR(dirname(outshape), layer = basename(outshape), verbose=!quiet)
    return(shp)
  }
  return(NULL)
}

sn.new.inf <- raster('data/out/new1000.tif')
old.outlines <- readOGR('data/in/shapefile','FinalWshed_All_Whole') %>% spTransform(.,projection(sn.new.inf))

sn.1000 <- crop(sn.new.inf,extent(old.outlines))
sn.1000[sn.1000 < 1] <- NA
sn.1000.shp <- gdal_polygonizeR(sn.1000)
sn.1000.clip <- sn.1000.shp[old.outlines,]
sn.1000.clip <- gUnaryUnion(sn.1000.clip, id = sn.1000.clip$DN)
sn.1000.clip$name <- 'data'
# writeOGR(sn.1000.clip,'data/in/shapefile','sn.1000.clip',driver='ESRI Shapefile',overwrite_layer = T)

old.inf <- raster('old100.tif')
old.inf[old.inf<1] <- NA
old.1000.shp <- gdal_polygonizeR(old.inf)
old.100.clip <- old.1000.shp[old.outlines,]
old.100.clip <- gUnaryUnion(old.100.clip,id=old.100.clip$DN)
old.100.clip$name <- 'data'
# writeOGR(old.100.clip,'data/in/shapefile','old.100.clip',driver='ESRI Shapefile')

```


### Reproject outlets to same projection as rasters

```{r, eval = F}
#Match outlet projections to raster data
outs <- readOGR('data/in/shapefile/SubSheds.kml','Sub Sheds',stringsAsFactors=F) %>% spTransform(.,CRS=CRS('+init=epsg:26917'))
outs <- outs[,1]
outs$Id <- 1:nrow(outs)
outs <- outs[,2:1]


writeOGR(outs,'data/in/shapefile','approxnew',driver='ESRI Shapefile',overwrite_layer = T)




```
